---
title: "Lab 6"
author: "Alain Perez"
date: "3/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(babynames)
library(tidyverse)
jessie<-filter(babynames, name=="Jessie")


```
Creating a plot to see the number of Jesses born female and male.

```{r}
jessie %>%
  ggplot(aes(x=year, y=n, color=sex))+geom_line()
  


```

##Excercise 2 ##

We cannot mutate because we already have a variable that keeps track of the sex of the baby we would want to count the number of babies born of each sex per year.

Trying out summarize

```{r}
jessie%>%
  filter(year == 1982)%>%
  summarize(total=sum(n),
            num_males=sum(ifelse(sex=="M", n, 0)))

```
## Excercise 3 ##
```{r}
jessie%>%
  filter(year == 1982)%>%
  summarize(total=sum(n),num_males=sum(ifelse(sex =="M",n,0)))%>%
  mutate(male_prop=num_males/total)
```

Slicing files by year:

```{r}
jessie_by_year<- babynames%>%
  filter(name =="Jessie")%>%
  group_by(year)%>%
  summarize(
    num_rows=n(),
    total_births=sum(n),
    num_males=sum(ifelse(sex=="M",n,0)),
    num_famles=sum(ifelse(sex=="F", n,0)))%>%
  mutate(prop_males=num_males/total_births)
jessie_by_year
    
```
##Excercise 4##
```{r}
Jessie_E4<-babynames%>%
  filter(name == "Jessie")%>%
  group_by(year)%>%
  summarize(total_births=sum(n))%>%
  arrange(desc(total_births))
  head(Jessie_E4)

```
The myear with the highest total number of Jessies born was 1918 and there were 5003 Jessies born.

##Excercise 5##

```{r}
babynames%>%
  group_by(name)%>%
  summarize(
    first_year = min(year),
    last_year = max(year)
  )
  
```

##Excercise 6##

```{r}
babynames%>%
  group_by(name)%>%
  summarize(num_rows =n())%>%
  filter(num_rows ==2*(2015-1880+1))

```

##Excercise 7##

```{r}
babynames%>%
  filter(year<2000 & year>1990)%>%
  group_by(name)%>%
  summarize(num_births =sum(n))%>%
  arrange(desc(num_births))
  
```

##Excercise 8##

```{r}
popular_names_1997<-
  babynames%>%
  filter(year ==1997)%>%
  group_by(name)%>%
  summarize(total_for_name=sum(n))%>%
  mutate(overall_total=sum(total_for_name))%>%
  mutate(proportion_of_total= total_for_name/overall_total)%>%
  filter(proportion_of_total>=.01)%>%
  select(name)
popular_names_1997

babynames%>%
  filter(name %in% popular_names_1997$name)%>%
  group_by(name)%>%
  summarize(first_year=min(year))%>%
  summarize(youngest_name= name[which.max(first_year)], first_appearance= max(first_year))
  

```
##Excercise 9##

```{r}
babynames%>%
  group_by(year)%>%
  summarize(
    total_names=n(),
    num_male_name=sum(ifelse(sex =="M",1,0)),
    num_female_name=sum(ifelse(sex=="F",1,0)))%>%
      ggplot(aes(x=year))+geom_line(aes(y=total_names),color="green")+geom_line(aes(y=num_male_name), color="blue")+geom_line(aes(y=num_female_name), color="red")+ylab("Number of different names")
    
  


```

##Excercise 10##

```{r}
babynames%>%
  filter(name=="Sam")%>%
  group_by(year)%>%
  summarize(
    total_Sams=n(),
    num_male_Sams=sum(ifelse(sex =="M",1,0)),
    num_female_Sams=sum(ifelse(sex=="F",1,0)))%>%
      ggplot(aes(x=year))+geom_line(aes(y=total_Sams),color="green")+geom_line(aes(y=num_male_Sams), color="blue")+geom_line(aes(y=num_female_Sams), color="red")+ylab("Total Sams born/year")
    


```

###LAB7###
```{r}
total_births<-total_births%>%
  select(-num_rows)%>%
  rename(ssa=births.x, census=births.y)
head(total_births)

```